Semantic Novelty Detection with Graphical Models

### 2011-05-05 semantic-nd/article.tex
Added several tikz graphs explaining distributions.
Though for lack of time from now on other drawing methods will be used
and the tikz images will be developed later.

% Additional packages
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{tikz}
\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
\hypersetup{%
   plainpages=false, 
   pdfpagelayout=SinglePage,
   bookmarksopen=false,
   bookmarksnumbered=true,
   breaklinks=true,
   linktocpage,
   colorlinks=true,
   linkcolor=blue,
   urlcolor=blue,
   citecolor=blue,
   anchorcolor=green
}

% TiKZ styles
\tikzstyle{var} = [circle,
                   thick,
                   draw, fill=red!20,
                   text width=4.5em,
                   text badly centered,
                   node distance=3cm,
                   inner sep=0pt]

\tikzstyle{factor} = [rectangle,
                      thick,
                      draw, fill=blue!20,
                      text width=1em,
                      text centered,
                      minimum height=3em,
                      minimum width=3em]

@@

\begin{figure}[h]
\caption{\label{fig:unsupervised-threshold}By producing graphical models $G$ and $G'$ to both conditional and unconditional probability
         distributions on a set of variables, novelty can be detect on the threshold $P_G(x)/P_{G'}(x)$.
         In this case the unconditional probability is modelled as a fully uniform distribution.}
\end{figure}

@@

\begin{figure}[h]
\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to create a better approximation
         for the unconditional probability. Which should lead to a better novelty threshold.}
\end{figure}

>>>>

% Additional packages
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{subfig}
\usepackage{tikz}
\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
\hypersetup{%
   plainpages=false, 
   pdfpagelayout=SinglePage,
   bookmarksopen=false,
   bookmarksnumbered=true,
   breaklinks=true,
   linktocpage,
   colorlinks=true,
   linkcolor=blue,
   urlcolor=blue,
   citecolor=blue,
   anchorcolor=green
}

% TiKZ styles
\tikzstyle{var} = [circle,
                   thick,
                   draw, fill=red!20,
                   font=\scriptsize,
                   text width=3.5em,
                   text centered,
                   node distance=1em,
                   inner sep=0pt]

\tikzstyle{factor} = [rectangle,
                      thick,
                      draw, fill=blue!20,
                      minimum height=2em,
                      minimum width=2em]

@@

\begin{figure}[h]
\centering
\subfloat[Factor graph $G$ used to model conditional probability]{
\begin{tikzpicture}
  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
    \& \node [var] (room) {Room Category}; \& \\

    \node [factor] (fAppearance) {}; \&
    \node [factor] (fShape) {}; \&
    \node [factor] (fSize) {}; \\

    \node [var] (appearance) {Appearance}; \&
    \node [var] (shape) {Shape}; \&
    \node [var] (size) {Size}; \\
  };
  \path [line] (room) -- (fAppearance);
  \path [line] (room) -- (fShape);
  \path [line] (room) -- (fSize);

  \path [line] (fAppearance) -- (appearance);
  \path [line] (fShape) -- (shape);
  \path [line] (fSize) -- (size);
\end{tikzpicture}}
\hfill
\subfloat[Factor graph $G'$ used to model a unconditional probability distribution]{
\begin{tikzpicture}
  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
    \& \node [var] (room) {Room Category}; \& \\

    \node [var] (appearance) {Appearance}; \&
    \node [var] (shape) {Shape}; \&
    \node [var] (size) {Size}; \\
  };
\end{tikzpicture}}

\caption{\label{fig:unsupervised-threshold}
         Assuming a uniform unconditional distribution of any subset of the variables $x$,
         a novelty threshold can be defined on the function $P_G(x)/P_{G'}(x)$.}
\end{figure}

@@

\begin{figure}[h]
\centering
\subfloat[Factor graph $G$ used to model conditional probability]{
\begin{tikzpicture}
  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
    \& \node [var] (room) {Room Category}; \& \\

    \node [factor] (fAppearance) {}; \&
    \node [factor] (fShape) {}; \&
    \node [factor] (fSize) {}; \\

    \node [var] (appearance) {Appearance}; \&
    \node [var] (shape) {Shape}; \&
    \node [var] (size) {Size}; \\
  };
  \path [line] (room) -- (fAppearance);
  \path [line] (room) -- (fShape);
  \path [line] (room) -- (fSize);

  \path [line] (fAppearance) -- (appearance);
  \path [line] (fShape) -- (shape);
  \path [line] (fSize) -- (size);
\end{tikzpicture}}
\hfill
\subfloat[Factor graph $G'$ used to model a unconditional probability distribution]{
\begin{tikzpicture}
  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
    \node [var] (appearance) {Appearance}; \&
    \node [var] (shape) {Shape}; \&
    \node [var] (size) {Size}; \\

    \node [factor] (fAppearance) {}; \&
    \node [factor] (fShape) {}; \&
    \node [factor] (fSize) {}; \\
  };
  \path [line] (fAppearance) -- (appearance);
  \path [line] (fShape) -- (shape);
  \path [line] (fSize) -- (size);
\end{tikzpicture}}

\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to create a better approximation
         for the unconditional probability. Which should lead to a better novelty threshold.}
\end{figure}
